Comparative study of requantization of time-dependent mean-field for pairing dynamics

### 2018-02-07 draft2.2.tex
Revision on 2018.2.7

The ground states in many of even-even nuclei are well
described by the Bardeen-Cooper-Schriefer (BCS)
and the Hartree-Fock-Bogoliubov (HFB) theories \cite{RS80,BHR03}.
Its time-dependent version, 
the time-dependent HFB (TDHFB) theory \cite{BR86,NMMY16},
is a natural extension of the static HFB theory.
The small-amplitude limit of the TDHFB is known to be 
the quasiparticle random phase approximation (QRPA).
%Because of the improvement of computational power,
%dynamics and reaction process in nuclei have been investigated
%based on QRPA.
The QRPA has been extensively utilized and successfully
describes properties of giant resonances.
In contrast, many of low-lying excited states
cannot be well reproduced by QRPA \cite{NMMY16}.
This may be due to their slowly moving large-amplitude nature.

@@

In principle, the TDHFB dynamics can be applicable to large amplitude motion.
The problem is that it is not easy to determine quantum mechanical 
quantities, such as energy eigenvalues and transition matrix elements,
from the TDHFB trajectories.
In addition, the TDHFB lacks a part of quantum fluctuation
associated with low-energy
large amplitude collective motion,
which leads to difficulties in description of the quantum tunneling
processes, such as spontaneous fission and sub-barrier fusion reaction \cite{S12}.
For this purpose,
since the TDHF(B) trajectory corresponds to a classical limit of the path
integral formulation,
the requantization of the mean-field dynamics was
proposed \cite{Neg82,L80,LNP80,KS80,K81,Rei80}.
It recovers quantum fluctuations missing in the mean-field
level, and possibly enables us to describe large-amplitude collective
tunneling phenomena such as spontaneous fission \cite{Neg82}.
The requantization of the TDHFB is particularly feasible
for integrable systems, because the system is described by separable
action-angle variables $(I_k, \phi_k)$, leading to
the Einstein-Brillouin-Keller (EBK) quantization condition.
However, for nonintegrable systems in general, it is difficult to find
suitable periodic orbits to quantize.
A possible solution to this difficulty is to find a decoupled collective
subspace spanned by
a single coordinate and its conjugate momentum \cite{NMMY16}.
Since the one-dimensional system is integrable, the quantization is
practicable.

@@

Starting an arbitrary state $\ket{\psi(0)}$ at time $t=0$,
the time-dependent full quantum state can be written in the path integral form
\begin{equation}
\ket{\psi(t)}
	= e^{-iHt} \ket{\psi(0)}
	= \int d\mu(Z'') \ket{Z''} \int d\mu(Z')
	\int_{Z(0)=Z'}^{Z(t)=Z''} \mathcal{D}\mu[Z(\tau)]
	e^{i\mathcal{S}[Z(\tau)]} \psi(Z') ,
\label{path_integral_expression}
\end{equation}
where $\psi(Z)\equiv \braket{Z|\psi(0)}$ and
the invariant measure $d\mu(Z)$ is defined by 
the unity condition,
\begin{equation}
  \int d\mu(Z) \ket{Z}\bra{Z} = 1 .
  \label{unity}
\end{equation}
In Eq. (\ref{path_integral_expression}),
$\mathcal{S}[Z(\tau)]$ is the action (\ref{S}) along a given path $Z(\tau)$
with the initial coherent state $\ket{Z(0)}=\ket{Z'}$ and
the final state $\ket{Z(t)}=\ket{Z''}$,
then, the integration $\int \mathcal {D}\mu[Z(\tau)]$
is performed over all possible paths $\ket{Z(\tau)}$ between them.
Among all trajectories in the path integral,
the lowest stationary-phase approximation selects
the TDHFB (classical) trajectories.
%\begin{align}
%	\int \mathcal{D}\mu(Z) e^{i\mathcal{S}(Z,Z_0)} \approx e^{i\mathcal{S}_{\rm cl}(Z,Z_0)} .
%\end{align}
\begin{equation}
	\ket{\psi(t)} \approx \int d\mu(Z') \ket{Z'_{\rm cl}(t)}
	e^{i\mathcal{S}_{\rm cl}(Z'_{\rm cl}(t),Z')} \psi(Z') ,
	\label{SPA}
\end{equation}
where the TDHFB trajectory starting from $\ket{Z'}$
ends at $\ket{Z'_{\rm cl}(t)}$ at time $t$.
The action $\mathcal{S}_{\rm cl}(Z_f,Z_i)$ is calculated along this
classical trajectory
connecting $Z_i=Z'_{\rm cl}(0)=Z'$ and $Z_f=Z'_{\rm cl}(t)$.
\begin{equation}
	\mathcal{S}_{\rm cl}(Z'_{\rm cl}(t),Z') \equiv \int_{0}^{t}
\braket{Z_{\rm cl}(t)|
	i\frac{\partial}{\partial t} - H
	|Z_{\rm cl}(t)} dt
	= \mathcal{T}[Z_{\rm cl}] - \mathcal{H}(Z',Z'^*) t 
	,
	\label{S_cl}
\end{equation}
with
\begin{equation}
\mathcal{T}[Z_{\rm cl}]\equiv
\int_{0}^{t} \braket{Z_{\rm cl}(t)| i\frac{\partial}{\partial t}
	|Z_{\rm cl}(t)} dt 
	= \int_{Z'}^{Z'_{\rm cl}(t)}
	\frac{i}{2} \sum_l \frac{\Omega_l}{1+|Z_l|^2}
  (Z_l^* dZ_l - Z_l dZ_l^*) .
\end{equation}
In the last equation of Eq. (\ref{S_cl}),
we used the fact that the TDHFB trajectory conserves the energy,
$\mathcal{H}(Z_{\rm cl}(t),Z_{\rm cl}^*(t))= \mathcal{H}(Z',Z'^*)$.
%We assume that only classical (TDHFB) trajectory contributes to the time dependent wave function

@@

\bibliographystyle{apsrev4-1}
\bibliography{ref}

>>>>

The ground states in many of even-even nuclei are well
described by the Bardeen-Cooper-Schriefer (BCS)
and the Hartree-Fock-Bogoliubov (HFB) theories \cite{RS80,BHR03}.
Its time-dependent version, 
the time-dependent HFB (TDHFB) theory \cite{BR86,NMMY16},
is a natural extension of the static HFB theory.
Thanks to increasing computational power,
realistic applications of the TDHFB calculations in real time
become available for studies of linear response properties
\cite{Eba10,SBMR11,Has12,SL13-2,ENI14}
and of various nuclear dynamics
\cite{SL13-1,SBBMR15,SSL15,WSMBF16,MSW17,HS17}.
The small-amplitude limit of the TDHFB is known to be 
the quasiparticle random phase approximation (QRPA).
%Because of the improvement of computational power,
%dynamics and reaction process in nuclei have been investigated
%based on QRPA.
The QRPA has been extensively utilized and successfully
describes properties of giant resonances.
Recently, the QRPA calculations with modern energy density functionals
for giant resonances in deformed nuclei have become available
\cite{Nes06,YG08,PG08,AKR09,Los10,TE10,YN11,YN13}.
In contrast, many of low-lying excited states
cannot be well reproduced by QRPA \cite{NMMY16}.
This may be due to their slowly moving large-amplitude nature.

@@

In principle, the TDHFB dynamics can be applicable to large amplitude motion.
The problem is that it is not easy to determine quantum mechanical 
quantities, such as energy eigenvalues and transition matrix elements,
from the TDHFB trajectories.
In addition, the TDHFB lacks a part of quantum fluctuation
associated with low-energy
large amplitude collective motion,
which leads to difficulties in description of the quantum tunneling
processes, such as spontaneous fission and sub-barrier fusion reaction.
% \cite{S12}.
For this purpose,
since the TDHF(B) trajectory corresponds to a stationary-phase limit of the path
integral formulation,
the requantization of the mean-field dynamics was
proposed \cite{Neg82,L80,LNP80,KS80,K81,Rei80}.
It recovers quantum fluctuations missing in the mean-field
level, and possibly enables us to describe large-amplitude collective
tunneling phenomena.
%such as spontaneous fission \cite{Neg82}.
The requantization of the TDHFB is particularly feasible
for integrable systems, because the system is described by separable
action-angle variables $(I_k, \phi_k)$, leading to
the Einstein-Brillouin-Keller (EBK) quantization condition.
However, for nonintegrable systems in general, it is difficult to find
suitable periodic orbits to quantize.
A possible solution to this difficulty is to find a decoupled collective
subspace spanned by
a single coordinate and its conjugate momentum \cite{NMMY16}.
Since the one-dimensional system is integrable, the quantization is
practicable.

@@

Starting an arbitrary state $\ket{\psi(0)}$ at time $t=0$,
the time-dependent full quantum state can be written in the path integral form
\begin{equation}
\ket{\psi(t)}
	= e^{-iHt} \ket{\psi(0)}
	= \int d\mu(Z'') \ket{Z''} \int d\mu(Z')
	\int_{Z(0)=Z'}^{Z(t)=Z''} \mathcal{D}\mu[Z(\tau)]
	e^{i\mathcal{S}[Z(\tau)]} \psi(Z') ,
\label{path_integral_expression}
\end{equation}
where $\psi(Z)\equiv \braket{Z|\psi(0)}$ and
the invariant measure $d\mu(Z)$ is defined by 
the unity condition,
\begin{equation}
  \int d\mu(Z) \ket{Z}\bra{Z} = 1 .
  \label{unity}
\end{equation}
In Eq. (\ref{path_integral_expression}),
$\mathcal{S}[Z(\tau)]$ is the action (\ref{S}) along a given path $Z(\tau)$
with the initial coherent state $\ket{Z(0)}=\ket{Z'}$ and
the final state $\ket{Z(t)}=\ket{Z''}$,
then, the integration $\int \mathcal {D}\mu[Z(\tau)]$
is performed over all possible paths $\ket{Z(\tau)}$ between them.
Among all trajectories in the path integral,
the lowest stationary-phase approximation selects
the TDHFB (classical) trajectories\footnote{
n this formulation, the stationary-phase approximation agrees with
the TDHF(B) trajectories, while that to the auxiliary-field path
integral of Refs. \cite{Neg82,L80}
leads to the TDH(B) without the Fock potentials.
}.
%\begin{align}
%	\int \mathcal{D}\mu(Z) e^{i\mathcal{S}(Z,Z_0)} \approx e^{i\mathcal{S}_{\rm cl}(Z,Z_0)} .
%\end{align}
\begin{equation}
	\ket{\psi(t)} \approx \int d\mu(Z') \ket{Z'_{\rm cl}(t)}
	e^{i\mathcal{S}_{\rm cl}(Z'_{\rm cl}(t),Z')} \psi(Z') ,
	\label{SPA}
\end{equation}
where the TDHFB trajectory starting from $\ket{Z'}$
ends at $\ket{Z'_{\rm cl}(t)}$ at time $t$.
The action $\mathcal{S}_{\rm cl}(Z_f,Z_i)$ is calculated along this
classical trajectory
connecting $Z_i=Z'_{\rm cl}(0)=Z'$ and $Z_f=Z'_{\rm cl}(t)$.
\begin{equation}
	\mathcal{S}_{\rm cl}(Z'_{\rm cl}(t),Z') \equiv \int_{0}^{t}
\braket{Z_{\rm cl}(t)|
	i\frac{\partial}{\partial t} - H
	|Z_{\rm cl}(t)} dt
	= \mathcal{T}[Z_{\rm cl}] - \mathcal{H}(Z',Z'^*) t 
	,
	\label{S_cl}
\end{equation}
with
\begin{equation}
\mathcal{T}[Z_{\rm cl}]\equiv
\int_{0}^{t} \braket{Z_{\rm cl}(t)| i\frac{\partial}{\partial t}
	|Z_{\rm cl}(t)} dt 
	= \int_{Z'}^{Z'_{\rm cl}(t)}
	\frac{i}{2} \sum_l \frac{\Omega_l}{1+|Z_l|^2}
  (Z_l^* dZ_l - Z_l dZ_l^*) .
\end{equation}
In the last equation of Eq. (\ref{S_cl}),
we used the fact that the TDHFB trajectory conserves the energy,
$\mathcal{H}(Z_{\rm cl}(t),Z_{\rm cl}^*(t))= \mathcal{H}(Z',Z'^*)$.
%We assume that only classical (TDHFB) trajectory contributes to the time dependent wave function

@@

\bibliographystyle{apsrev4-1}
%\bibliography{ref}
\bibliography{ref2}
